---
title: "Assignment 3"
format:
  html:
    embed-resources: true
---

## Problem 1

### a.

``` stata
*read file in stata
. import sasxport5 "DEMO_D.XPT", clear

. save "DEMO_D.dta", replace
(file DEMO_D.dta not found)
file DEMO_D.dta saved
```

``` stata
. use "VIX_D.dta", clear

. merge 1:1 seqn using "DEMO_D.dta"

    Result                      Number of obs
    -----------------------------------------
    Not matched                         3,368
        from master                         0  (_merge==1)
        from using                      3,368  (_merge==2)

    Matched                             6,980  (_merge==3)
    -----------------------------------------

. keep if _merge==3
(3,368 observations deleted)

. count
  6,980
```

### b.

``` stata
* Generate age brackets
. egen age_bracket = cut(ridageyr), at(0(10)90) label

* generate a table displaying the proportions for "1" (Yes), "2" (No), "9" (Don't know)
. tabulate age_bracket viq220, row nofreq

           | Glasses/contact lenses worn for
age_bracke |             distance
         t |         1          2          9 |     Total
-----------+---------------------------------+----------
       10- |     32.09      67.91       0.00 |    100.00 
       20- |     32.59      67.20       0.21 |    100.00 
       30- |     35.87      64.13       0.00 |    100.00 
       40- |     37.00      63.00       0.00 |    100.00 
       50- |     55.01      44.99       0.00 |    100.00 
       60- |     62.22      37.78       0.00 |    100.00 
       70- |     66.89      33.11       0.00 |    100.00 
       80- |     66.88      33.12       0.00 |    100.00 
-----------+---------------------------------+----------
     Total |     42.23      57.74       0.03 |    100.00 
```

### c.

``` stata
* replace value 2 to 0 because in logistic regression, 0 = negative outcome,all other nonmissing values = positive outcome
. replace viq220 = 0 if viq220 ==2
(3,780 real changes made)
. replace viq220 = . if viq220 ==9
(2 real changes made, 2 to missing)
```

``` stata
* Fit three models and calculate AIC 
. logit viq220 ridageyr,or

Iteration 0:  Log likelihood = -4457.6265  
Iteration 1:  Log likelihood = -4236.2351  
Iteration 2:  Log likelihood = -4235.9433  
Iteration 3:  Log likelihood = -4235.9433  

Logistic regression                                     Number of obs =  6,545
                                                        LR chi2(1)    = 443.37
                                                        Prob > chi2   = 0.0000
Log likelihood = -4235.9433                             Pseudo R2     = 0.0497

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |    1.02498   .0012356    20.47   0.000     1.022561    1.027405
       _cons |    .283379   .0151461   -23.59   0.000     .2551952    .3146755
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,545  -4457.627  -4235.943       2   8475.887    8489.46
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

. estimates store model1

. logit viq220 ridageyr i.ridreth1 i.riagendr, or

Iteration 0:  Log likelihood = -4457.6265  
Iteration 1:  Log likelihood = -4138.3859  
Iteration 2:  Log likelihood = -4136.8807  
Iteration 3:  Log likelihood = -4136.8805  

Logistic regression                                     Number of obs =  6,545
                                                        LR chi2(6)    = 641.49
                                                        Prob > chi2   = 0.0000
Log likelihood = -4136.8805                             Pseudo R2     = 0.0720

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   1.022831   .0012912    17.88   0.000     1.020303    1.025365
             |
    ridreth1 |
          2  |   1.169203    .192081     0.95   0.341     .8473273    1.613349
          3  |   1.952149   .1366952     9.55   0.000     1.701803    2.239322
          4  |    1.29936   .0995052     3.42   0.001     1.118264    1.509783
          5  |   1.917442   .2596352     4.81   0.000     1.470495    2.500236
             |
  2.riagendr |    1.65217   .0875831     9.47   0.000     1.489127    1.833064
       _cons |   .1593479   .0124169   -23.57   0.000     .1367784    .1856414
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,545  -4457.627   -4136.88       7   8287.761   8335.266
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

. estimates store model2

. logit viq220 ridageyr i.ridreth1 i.riagendr indfmpir, or

Iteration 0:  Log likelihood = -4259.5533  
Iteration 1:  Log likelihood = -3948.3256  
Iteration 2:  Log likelihood = -3946.9043  
Iteration 3:  Log likelihood = -3946.9041  

Logistic regression                                     Number of obs =  6,247
                                                        LR chi2(7)    = 625.30
                                                        Prob > chi2   = 0.0000
Log likelihood = -3946.9041                             Pseudo R2     = 0.0734

------------------------------------------------------------------------------
      viq220 | Odds ratio   Std. err.      z    P>|z|     [95% conf. interval]
-------------+----------------------------------------------------------------
    ridageyr |   1.022436    .001324    17.14   0.000     1.019845    1.025035
             |
    ridreth1 |
          2  |   1.123021   .1889653     0.69   0.490     .8075333    1.561764
          3  |   1.651244   .1240886     6.67   0.000     1.425098    1.913277
          4  |   1.230456   .0974736     2.62   0.009     1.053503     1.43713
          5  |   1.703572   .2387583     3.80   0.000     1.294384    2.242114
             |
  2.riagendr |   1.675767   .0910025     9.51   0.000      1.50657    1.863967
    indfmpir |   1.120301   .0198376     6.42   0.000     1.082087    1.159865
       _cons |   .1331659   .0116903   -22.97   0.000     .1121161    .1581678
------------------------------------------------------------------------------
Note: _cons estimates baseline odds.

. estat ic

Akaike's information criterion and Bayesian information criterion

-----------------------------------------------------------------------------
       Model |          N   ll(null)  ll(model)      df        AIC        BIC
-------------+---------------------------------------------------------------
           . |      6,247  -4259.553  -3946.904       8   7909.808   7963.727
-----------------------------------------------------------------------------
Note: BIC uses N = number of observations. See [R] IC note.

. estimates store model3
```

``` stata
* Create a table contains results required
. esttab model1 model2 model3, eform cells(b(fmt(4))) stats(N r2_p aic, labels("Sample Size"  "Pseudo R^2" "AIC"))
>  mtitle("Model 1" "Model 2" "Model 3") label

-----------------------------------------------------------
                              (1)          (2)          (3)
                          Model 1      Model 2      Model 3
                                b            b            b
-----------------------------------------------------------
Glasses/contact le~                                        
Age at Screening A~R       1.0250       1.0228       1.0224
Race/Ethnicity - R~1                    1.0000       1.0000
Race/Ethnicity - R~2                    1.1692       1.1230
Race/Ethnicity - R~3                    1.9521       1.6512
Race/Ethnicity - R~4                    1.2994       1.2305
Race/Ethnicity - R~5                    1.9174       1.7036
Gender=1                                1.0000       1.0000
Gender=2                                1.6522       1.6758
Family PIR                                           1.1203
-----------------------------------------------------------
Sample Size             6545.0000    6545.0000    6247.0000
Pseudo R^2                 0.0497       0.0720       0.0734
AIC                     8475.8866    8287.7609    7909.8082
-----------------------------------------------------------
Exponentiated coefficients
```

### d.

``` stata
* Odds:
* The odd of men is 1, which is set as the base in this case. Odds equals p(success)/p(failure) = 1. Therefore the probability of men wearing glasess/contact lenses for distance vision or not is equal. In other words, men have 50% probability wearing glasses/contact.
* The odds of women is 1.6758. Odds equals p(success)/(1-p(success)) = 1.6758. Therefore the probability of women wearing glasess/contact lenses for distance vision is 62.6%.
* Odds Ratio = 1/1.6758 = 0.597. Therefore, for a male, the odds of wearing glasess/contact lenses for distance vision are 0.597 times as small as the odds for a female wearing glasess/contact lenses for distance vision. In other words, for a female, the odds of wearing glasess/contact lenses for distance vision are 1.6758 times as large as the odds for a male wearing glasess/contact lenses for distance vision

* Proportion:
. tabulate riagendr viq220, row

+----------------+
| Key            |
|----------------|
|   frequency    |
| row percentage |
+----------------+

           |    Glasses/contact
           |    lenses worn for
           |       distance
    Gender |         0          1 |     Total
-----------+----------------------+----------
         1 |     2,014      1,181 |     3,195 
           |     63.04      36.96 |    100.00 
-----------+----------------------+----------
         2 |     1,766      1,584 |     3,350 
           |     52.72      47.28 |    100.00 
-----------+----------------------+----------
     Total |     3,780      2,765 |     6,545 
           |     57.75      42.25 |    100.00 
 

* From the table above, 
  36.96% of male wear glasses/contact lenses for distance vision,
  63.04% of male do not wear glasses/contact lenses for distance vision,
  47.28% of female wear glasses/contact lenses for distance vision,
  52.72% of female do not wear glasses/contact lenses for distance vision,
  Overall, there are 42.71% of male and 57.29% female of wearers of glasses/contact lenses for distance vision.

*In conclusion, male is less likely to wear glasses/contact lenses for distance vision than women. the result of proportion is the same as the result of odds ratio.
```

## Problem 2

```{r}
# Load the data into a data frame using RSQLite
library(DBI)
sakila <- dbConnect(RSQLite::SQLite(), "sakila_master.db")
dbListTables(sakila)
gg <- function(x) {
  dbGetQuery(sakila, x)
}
```

### a.

#### SQL

```{r}
gg("
SELECT language.name, COUNT(*) AS count
FROM film
JOIN language ON film.language_id = language.language_id
GROUP BY language.name
ORDER BY count DESC
")
```

As we can see from the result, there is no language which is most common for films aside from English.

### b.

#### SQL

```{r}
gg("
SELECT category.name, COUNT(*) AS count
FROM film
JOIN film_category ON film.film_id = film_category.film_id
JOIN category ON film_category.category_id = category.category_id
GROUP BY category.name
ORDER BY count DESC
LIMIT 1;
")
```

#### R

```{r}
films <- gg("SELECT * FROM film_category")
category <- gg("SELECT * FROM category")
# Find the most common genre
genre_counts <- table(films$category_id)
top_genreID <- names(genre_counts)[which.max(genre_counts)]

# Find the genre name of most common genre id
top_genrename <- category$name[category$category_id == top_genreID]
genre_counts
top_genreID
top_genrename
```

Sports is the most common genre in the data, and there are 74 movies of this genre.

### c.

#### SQL

```{r}
gg("
SELECT country, COUNT(*) AS count
FROM customer_list
GROUP BY country
HAVING count = 9
")
```

#### R

```{r}
customer <- gg("SELECT * FROM customer_list")
country_counts <- table(customer$country)
customers9 <- names(country_counts[country_counts == 9])
customers9
```

United Kingdom has exactly 9 customers.

## Problem 3

```{r}
us500 <- read.csv("us-500.csv")
```

```{r}
library(dplyr)
```

### a.

```{r}
prop1 <- us500 %>%
  filter(grepl("@[[:alnum:]_.]+\\.net", email)) %>%
  nrow() / nrow(us500)
prop1
```

The proportion of email addresses are hosted at a domain with TLD \".net" is 14%

### b.

```{r}
prop2 <- us500 %>%
  filter(grepl("[^[:alnum:]@.]", email)) %>%
  nrow() / nrow(us500)
prop2
```

The proportion of email addresses have at least one non alphanumeric character in them is 24.8%

### c.

```{r}
phone <- c(us500$phone1, us500$phone2)
area_code <- substr(phone, 1, 3)
codes_count <- table(area_code)
which.max(codes_count)
```

The most common area code amongst all phone numbers is 973.

### d.

```{r}
apartment_num <- gsub(".*#(\\d+)", "\\1", us500$address)
number = as.numeric(apartment_num)
hist(log(number), main = "Log of Apartment Numbers", xlab = "Log(Apartment Number)")

```

### e.

```{r}
digits <- substr(apartment_num, 1, 1)
digits_counts <- table(digits)
#apartment number proportion in data
actual_prob <- digits_counts/sum(digits_counts)
actual_prob
```

The Benford's law of proportions of leading apartment numbers are 0.301, 0.176, 0.125, 0.097, 0.079, 0.067,0.058, 0.051, 0.046, which is decreasingly distributed.

But the actual proportions of leading apartment numbers are 0.112, 0.120, 0.122, 0.114, 0.096, 0.110, 0.114, 0.090, 0.122, which is equally distributed.

It is obviously that the apartment numbers do not follow the Benford's law and the apartment numbers would not pass as real data.

### f.

```{r}
street_num <- as.numeric(sub("^(.*?)\\s.*", "\\1", apartment_num))
last_digits <- street_num %% 10
last_counts <- table(last_digits)
last_prob <- last_counts/sum(last_counts)
last_prob
```

The Benford's law of proportions of last numbers are nearly 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, which is equally distributed.

The actual proportions of last street numbers are 0.084 0.088 0.128 0.110 0.104 0.092 0.080 0.096 0.080 0.138, which is nearly equally distributed.

Therefore the last street numbers follow the Benford's law and the street numbers would pass as real data.
